---
title: "Reproducible Research: Peer Assessment 1"
author: "GJAllen"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

### Loading and preprocessing the data
Just reading in the data, but let's also require some packages that I like.

```{r}
library(data.table)
library(dplyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(magrittr, warn.conflicts = FALSE)
activity <- data.table::fread("activity.csv") %>% 
  mutate(date = as.Date(date))
head(activity)
```

### What is mean total number of steps taken per day?
Let's first make a histogram:
```{r}
steps.by.day <- activity %>% 
  filter(!is.na(steps)) %>% 
  group_by(date) %>% 
  summarize(steps = sum(steps, na.rm = TRUE))
  
hist(steps.by.day$steps, breaks = 10, xlab = "Total Steps per Day", main = "Total Steps per Day")
```

The mean:
```{r}
mean(steps.by.day$steps, na.rm = TRUE)
```

The median:
```{r}
median(steps.by.day$steps, na.rm = TRUE)
```

### What is the average daily activity pattern?
Let's have a look first at the pattern (sorry I'm really stretching).
```{r}
interval.aggregate <- activity %>% 
  group_by(interval) %>% 
  summarize(steps = mean(steps, na.rm = TRUE))
ggplot(interval.aggregate) +
  aes(interval, steps) +
  geom_line()
```

And now for the maximum (it's interval 835):
```{r}
filter(interval.aggregate, steps == max(steps))
```

### Imputing missing values
The total number of missing values (it looks like there are 2304 missing values for steps):
```{r}
colSums(is.na(activity))
```

Let's impute the missing values. In this case, we'll just use the average value for each interval.
```{r}
imputed.activity <- interval.aggregate %>% 
  rename(average.steps = steps) %>% 
  right_join(activity, "interval") %>% 
  mutate(steps = ifelse(is.na(steps), average.steps, steps)) %>% 
  select(-average.steps)
```

Now the histogram, mean, and median:
```{r}
steps.by.day.imputed <- imputed.activity %>% 
  group_by(date) %>% 
  summarize(steps = sum(steps, na.rm = TRUE))

hist(steps.by.day.imputed$steps, breaks = 10, xlab = "Total Steps per Day", 
     main = "Total Steps per Day")
mean(steps.by.day.imputed$steps)
median(steps.by.day.imputed$steps)
```

There are differences.

### Are there differences in activity patterns between weekdays and weekends?
Let's add the weekday indicator:
```{r}
imputed.activity %<>% 
  mutate(weekday = weekdays(date),
         weekday = ifelse(weekday %in% c("Saturday", "Sunday"), 
                          "Weekend",
                          "Weekday"))
5/7
prop.table(table(imputed.activity$weekday))
```
It looks pretty reasonable.  Now lets add make the panel plot:
```{r}
imputed.activity %>%
  group_by(weekday, interval) %>% 
  summarize(steps = mean(steps, na.rm = TRUE)) %>% 
  ggplot() +
  aes(interval, steps) +
  geom_line() +
  facet_wrap(~ weekday)
```  
  
Yes there are differences.
